\emph{icepack}: a new glacier flow modeling package in Python, version 1.0

### 2021-04-06 reviews/response-3.tex

We do not claim to have reduced the need for expertise in glaciology itself, i.e. understanding mass and energy balance, etc.
We also consider having some understanding of the different boundary conditions that might apply to the system to be part of this knowledge.
The text has been expanded to describe these boundary conditions in more detail (see comments that follow).
Granted, (1) there is more demonstration work that we need to do in order to cover use cases like modeling a small mountain glacier and (2) there are technical hurdles we still need to overcome to do a better job handling the margins.
Very likely your recent work on variational inequalities will factor into improvement on the second front.

We have altered the text to state more clearly that our goal is to help users who are not experts in scientific computing.
We also added the following sentence to the conclusion: ``Exactly how to apply principles from HCI to maximize usability is, nonetheless, not an exact science.
Our implementation may have failed to meet this goal and changes in future versions will be guided by what users find most difficult.''

@@

\textcolor{blue}{300--301, C:  I seriously doubt that more than one user in 100 will want to substitute their own melt-fraction dependence, because no one has the data for it.  So: "likely" --> "possible".  (This is a clear case where good defaults are more important than modularity.)}

Changed to ``Some users of this module may want to substitute in their own parameterization for melt fraction dependence.''

\textcolor{blue}{297--305, C:  This paragraph reads, to me, like an argument that expert knowledge *will* be required for the Icepack user.  ("users must calculate the fluidity field themselves")  That is, unless Icepack includes both defaults and higher-level simulation drivers which allow the novice user to never know about these possibilities.  More broadly, what does Icepack design offer, when it comes to avoiding explicit choices of all submodels, on the way to a first scientifically-useful model for the beginner?}

What you're describing is knowledge of glaciology, not scientific computing.
We have stated explicitly in the first sentence of section 6 that we aim to reduce the need for users to be experts in the latter but not the former.
If we were to define how some fields are coupled, we would be faced with a difficult design choice of where this is or is not appropriate.
Including the effect of temperature on rheology seems fairly obvious.
Some modeling papers also account for much higher fluidity for ice that accumulated during the last glacial period because of the greater dust content.
Should it also be the responsibility of the package to include this effect?
We did not feel that, at this stage of the development, we could also implement high-level simulation drivers that hide these details while also satisfying our other design goals.
See also response to the first comment and the text added to the conclusion.
We believe that our design choices will ultimately enhance usability for the intended audience and we did our best by applying principles from HCI.
But we may well have missed the mark and the interface may change to respond to evolving understanding of our target audience's needs.

@@

\textcolor{blue}{474--487:  Extruded meshes.  On board with Firedrake.  But if the model claims to have a solution of (1) then there would be some mechanism for addressing regions which become ice-free within a time step?  No mention thereof.}

You can use extruded meshes of variable thickness and that thickness can become 0.
So the choice of extruded meshes does not force one to use terrain-following coordinates as we did, whether or not that was a good idea.
No change to the text.

\textcolor{blue}{493--515, A:  Section 4.5 describes a nice solution to a real problem.  In fact, can one consider this kind of fit to the glacier profile at a grounded margin?  Not obvious how to proceed, I agree, but note that the thickness field h of a glacier has the similar continuous-but-non-differentiable character as the pressure on the ocean calving front.}

The technique might be applicable in other areas but we haven't had the chance to try it yet.
No change to the text.

### 2021-04-06 icepack.tex
Responses to 3

We have also paid special attention to how we can design this software package to be most usable for its intended audience.
Relatively few works in the computational science literature draw directly from relevant work in HCI when discussing usability; see for example \citet{hannay2009scientists, harris2020array}.
We believe that this is because of two difficulties.
First, the degree to which usability is a rate-limiting factor for scientists is hard to quantify and likely differs widely across disciplines.
Second, concretely assessing what features make software tools more or less usable is highly subjective.
By contrast, measuring computational performance is much more feasible although still fraught with difficulties of its own.
(This is not to say that performance optimization is easy by any means.)
In working with several graduate students and postdoctoral researchers in glaciology, we have observed that usability is a substantial bottleneck for researchers at these career stages.
For this reason, we have chosen to focus explicitly on usability by drawing on the research literature in HCI.
We argue that the same design features that enhance usability for relative novices to the subject area will also enhance the productivity of expert users.

This paper has presented several physics models currently implemented in icepack along with demonstrations.
Future developments will include:
\begin{enumerate}
    \item an implementation of full Stokes flow,
    \item improved physics formulations and solvers that work in ice-free areas,
    \item adaptivity in time and space,
\end{enumerate}
New features will be guided by feedback from icepack users and from the glaciological community at large.
By providing implementations of several glacier flow models from less to more complex and by enabling users to experiment with the physics, this tool both lowers the barrier to entry for novices to glacier flow modeling and provides a pathway for these users to progress towards ever more sophisticated and advanced simulations.

>>>>

We have also paid special attention to how we can design this software package to be most usable for its intended audience.
Relatively few works in the computational science literature draw directly from relevant work in HCI when discussing usability; see for example \citet{hannay2009scientists, harris2020array}.
We believe that this is because of two difficulties.
First, the degree to which usability is a rate-limiting factor for scientists is hard to quantify and likely differs widely across disciplines.
Second, concretely assessing what features make software tools more or less usable is highly subjective.
By contrast, measuring computational performance is much more feasible although still fraught with difficulties of its own.
(This is not to say that performance optimization is easy by any means.)
In working with several graduate students and postdoctoral researchers in glaciology, we have observed that usability is a substantial bottleneck for researchers at these career stages.
For this reason, we have chosen to focus explicitly on usability by drawing on the research literature in HCI.
Exactly how to apply principles from HCI to maximize usability is, nonetheless, not an exact science.
Our implementation may have failed to meet this goal and changes in future versions will be guided by what users find most difficult.
Finally, we argue that the same design features that enhance usability for relative novices to the subject area will also enhance the productivity of expert users.

This paper has presented several physics models currently implemented in icepack along with demonstrations.
Future developments will include:
\begin{enumerate}
    \item an implementation of full Stokes flow,
    \item improved physics formulations and solvers that work in ice-free areas,
    \item adaptivity in time and space,
\end{enumerate}
New features will be guided by feedback from icepack users and from the glaciological community at large.
By providing implementations of several glacier flow models from less to more complex and by enabling users to experiment with the physics, this tool both lowers the barrier to entry for novices to numerical modeling and provides a pathway for these users to progress towards ever more sophisticated and advanced simulations.
